Miniproject - Elliptic Curves

### 2022-03-31 Elliptic Curves/submission.tex
Remade example 1.7

\begin{example}[1(i)]
    Have
    \begin{equation*}
        |463^2 + 6|_5 = |214375|_5 = |5^4 \cdot 7^3|_5 = 5^{-4} < 5^{-3}
    \end{equation*}
    To find it, note that $|2^2 + 6|_5 < 1$ and use Newton's method.
    Set $x_0 = 2$ and have
    \begin{align*}
        x_1 = x_0 - \frac {x_0^2 + 6} {2x_0} = 2 - \frac {10} {4} = -\frac 1 2  \\
        x_2 = -\frac 1 2 - \frac {1/4 + 6} {-1} = \frac {25} 4 - \frac 1 2 = \frac {23} 4
    \end{align*}
    and indeed, $|(23/4)^2 + 6|_5 = |625/16|_5 = 5^{-4}$.
    Since the valuation $|\cdot|_5$ is non-Archimedean, observe that $|x^2 + 6|_5 < 5^{-3}$ holds for all $x \in \Q$ with $|x - 32/4|_5 = |4x - 32|_5 < 5^{-3}$.
    Hence, we look for $x \in \Z$ such that $5^4 \divides 4x + 23$.
    In other words, find $k \in \Z$ with $4 \divides k 5^4 - 23$, i.e. $k - 3 \equiv 0 \mod 4$.
    We find $k = 3$ and so $x = 463$.
\end{example}
\begin{example}[1(ii)]
    Let
    \begin{equation*}
        \alpha = 5^{-1} + 2 \cdot 5^0 + 5 (1 + 4 \cdot 5) \sum_{n \geq 0} 5^{2n} \in \Q_5
    \end{equation*}
    Note that in $\Q_5$ we have
    \begin{equation*}
        \sum_{n \geq 0} 5^{2n} = \sum_{n \geq 0} 25^n = \frac 1 {1 - 25} = -\frac 1 {24}
    \end{equation*}
    So
    \begin{equation*}
        \alpha = \frac 1 5 + 2 + 5 (21) \frac 1 {24} = \frac {263} {40}
    \end{equation*}
\end{example}
For the next exercises, we will slightly abuse notation and write
\begin{equation*}
    E(R) := \{ (x, y) \in E(\bar{k}) \ | \ x, y \in R \} \cup \{\O\}
\end{equation*}
for an Elliptic Curve $E$ defined over $k$ and any ring $R$ contained in the algebraic closure $\bar{k}$.
Note that this is in general not a group.
However, it gives us a way to nicely phrase the next statements.
\begin{prop}[1(iii)]
    \label{prop:1iii}
    Consider the Elliptic Curve $E: y^2 = x^3 + 2x + 2$ defined over $\Z$.
    Then $E(\Z) = \{ \O \}$ and
    \begin{equation*}
        E(\Z_p) \neq \{ \O \} \ \Leftrightarrow \ p \neq 3
    \end{equation*}
\end{prop}
\begin{proof}
    First show that there exists some $(x, y) \in \tilde{E}(\F_p)$ with $y \neq 0$ for all primes $p \neq 3$.

@@

    Now consider any prime $p \neq 2, 3$ and $(\tilde{x}, \tilde{y}) \in \tilde{E}(\F_p), x, y \in \Z, \tilde{y} \neq 0$.
    Let $f(t) := t^2 - x^3 - 2x - 2$.
    Then
    \begin{equation*}
        |f(y)|_p \leq p^{-1} \quad \text{and} \quad |f'(y)|_p = |y|_p = 1
    \end{equation*}
    Thus $|f(y)|_p < |f'(y)|_p^2$ and Hensel's Lemma yields a root $\gamma \in \Z_p$ with $(x, \gamma) \in E(\Z_p)$.

@@

    For the second part, note that by assumption, there is $(\tilde{x}, \tilde{y}) \in \tilde{E}(\F_p), x, y \in \Z$ with $|y|_p = 1$ and so
    \begin{equation*}
        |y^2 - x^3 - a_2 x^2 - a_4 x - a_6|_p \leq p^{-1} < 1 = 1^2 = |2y|_p
    \end{equation*}
    Hensel's Lemma now \cite[Thm 2.14]{lecture} shows that there is $\gamma \in \Z_p$ such that $\gamma^2 = x^3 + a_2 x^2 + a_4 x + a_6$ and so $(x, \gamma) \in E(\Z_p)$.
    Since $|y|_p = 1$ clearly also $|\gamma|_p = 1$.
    The other direction is obvious and follows directly by taking the reduction modulo $p$.

@@

    So assume now that $p \divides \Delta(E)$, hence $x^3 + f_2 x^2 + f_1 x + f_0$ factors as
    \begin{equation*}
        x^3 + \tilde{f}_2 x^2 + \tilde{f}_1 x + \tilde{f}_0 \equiv (x - \alpha)^2(x - \beta)
    \end{equation*}
    with $\alpha, \beta \in \bar{\F}_p$.
    However, note that $\F_p$ is perfect, so $(x - \alpha)^2(x - \beta)$ cannot be irreducible over $\F_p$, otherwise $\F_p[x] / \langle(x - \alpha)^2(x - \beta)\rangle$ would be a non-separable field extension of $\F_p$.
    Thus, either $\alpha \in \F_p$ or $\beta \in \F_p$.
    If $\alpha \in \F_p$, then clearly also $\beta = -2\alpha - \tilde{f}_2 \in \F_p$.
    If $\beta \in \F_p$, then also $(x - \alpha)^2 \in \F_p[x]$ and again by perfectness of $\F_p$, note that $\alpha \in \F_p$.
    So $\alpha, \beta \in \F_p$.

@@

    Interestingly, if we want a curve $E$ such that $\tilde{E}(\F_3) = \{ \O \}$, then we require
    \begin{equation*}
        f(0) = B \equiv 2  \quad \text{and} \quad f(1) = 1 + A + B \equiv 2
    \end{equation*}
    and this has the only solution $A \equiv B \equiv 2$, which indeed works out (this is the curve from Proposition~\ref{prop:1iii}).
\end{example}
This also works for $p = 7$, as the next example shows.
\begin{example}[1(iv) - second example]
    Assume we want to find a curve $E: y^2 = f(x) := x^3 + a_2 x^2 + a_4 x + a_6$ such that $\tilde{E}(\F_7)$ contains no point $(x, y)$ with $y \neq 0$.
    By considering a scaling isomorphism of the form
    \begin{equation*}
        E \to E', \quad (x, y) \mapsto (u^2 x, u^3 y)
    \end{equation*}
    we can assume wlog that $a_2 \equiv 0, 1, 2 \mod 7$ (as each element in $\F_7$ is congruent to $0, 1$ or $2$ modulo the squares $\{ x^2 \ | \ x \in \F_7 \}$).
    Furthermore have
    \begin{equation*}
        f(0) = a_6 \equiv 0, 3, 5, 6, \quad f(1) = 1 + a_2 + a_4 + a_6 \equiv 0, 3, 5, 6
    \end{equation*}
    since $1, 4, 2$ are the squares in $\F_7^*$.
    We find 48 possible solutions, and for example $a_2 \equiv 2, a_4 \equiv 6, a_6 \equiv 5$ works.
    Indeed, have that
    \begin{equation*}
        \tilde{E}(\F_7) = \{ (1, 0), (5, 0), (6, 0), \O \}
    \end{equation*}
    Furthermore, note that
    \begin{align*}
        1^2 &\equiv 2^3 + 2 \cdot 2^2 + 6 \cdot 2 + 5 = 33 \mod 8 \\
        1^2 &\equiv 2^3 + 2 \cdot 2^2 + 6 \cdot 2 + 5 = 33 \mod 3 \\
        2^2 &\equiv 1^3 + 2 \cdot 1^2 + 6 \cdot 1 + 5 = 14 \mod 5
    \end{align*}
    so Hensel's Lemma yields \cite[Thm 1.15]{lecture} points $(x, y) \in E(\Z_2)$ with $|y|_2 = 1$ and $(x', y') \in E(\Z_3)$ with $|y'|_3 = 1$ and $(x'', y'') \in E(\Z_5)$ with $|y''|_5 = 1$.
\end{example}
\begin{example}[(iv) - third example]
    The cubic polynomial $f = x^3 + x$ has roots $0$ and $1$ modulo 2.
    Thus the Elliptic Curve $E: y^2 = x^3 + x$ clearly has no points $(x, y) \in E(\Z_2)$ with $|y|_2 = 1$, as all points $(x, y) \in \tilde{E}(\F_2)$ satisfy $y = 0$.
    Furthermore, see that
    \begin{equation*}
        \tilde{E}(\F_5) = \{ (0, 0), (2, 0), (3, 0), \O \}
    \end{equation*}
    and
    \begin{align*}
        1^2 &\equiv 2^3 + 2 = 10 \mod 3 \\
        3^2 &\equiv 1^3 + 1 = 2 \mod 7
    \end{align*}
    so Hensel's Lemma yields \cite[Thm 1.15]{lecture} points $(x, y) \in E(\Z_3)$ with $|y|_3 = 1$ and $(x', y') \in E(\Z_7)$ with $|y'|_7 = 1$.
\end{example}
Note also that those three examples show that the bound $p \geq 11$ in Proposition~\ref{prop:points_in_EZp} is sharp.

>>>>

\begin{example}[1(i)]
    Have
    \begin{equation*}
        |463^2 + 6|_5 = |214375|_5 = |5^4 \cdot 7^3|_5 = 5^{-4} < 5^{-3}
    \end{equation*}
    To find it, note that $|2^2 + 6|_5 < 1$ and use Newton's method.
    Set $x_0 = 2$ and have
    \begin{align*}
        x_1 = x_0 - \frac {x_0^2 + 6} {2x_0} = 2 - \frac {10} {4} = -\frac 1 2  \\
        x_2 = -\frac 1 2 - \frac {1/4 + 6} {-1} = \frac {25} 4 - \frac 1 2 = \frac {23} 4
    \end{align*}
    and indeed, $|(23/4)^2 + 6|_5 = |625/16|_5 = 5^{-4}$.
    Since the valuation $|\cdot|_5$ is non-Archimedean, observe that $|x^2 + 6|_5 < 5^{-3}$ holds for all $x \in \Q$ with $|x - 32/4|_5 = |4x - 32|_5 < 5^{-3}$.
    Hence, we look for $x \in \Z$ such that $5^4 \divides 4x + 23$.
    In other words, find $k \in \Z$ with $4 \divides k 5^4 - 23$, i.e. $k - 3 \equiv 0 \mod 4$.
    We find $k = 3$ and so $x = 463$.
\end{example}
\begin{example}[1(ii)]
    Let
    \begin{equation*}
        \alpha = 5^{-1} + 2 \cdot 5^0 + 5 (1 + 4 \cdot 5) \sum_{n \geq 0} 5^{2n} \in \Q_5
    \end{equation*}
    Note that in $\Q_5$ we have
    \begin{equation*}
        \sum_{n \geq 0} 5^{2n} = \sum_{n \geq 0} 25^n = \frac 1 {1 - 25} = -\frac 1 {24}
    \end{equation*}
    So
    \begin{equation*}
        \alpha = \frac 1 5 + 2 + 5 (21) \frac 1 {24} = \frac {263} {40}
    \end{equation*}
\end{example}
For the next exercises, we will slightly abuse notation and write
\begin{equation*}
    E(R) := \{ (x, y) \in E(\bar{k}) \ | \ x, y \in R \} \cup \{\O\}
\end{equation*}
for an Elliptic Curve $E$ defined over $k$ and any ring $R$ contained in the algebraic closure $\bar{k}$.
Note that in general this is not a group.
However, it gives us a way to nicely phrase the next statements.
\begin{prop}[1(iii)]
    \label{prop:1iii}
    Consider the Elliptic Curve $E: y^2 = x^3 + 2x + 2$ defined over $\Z$.
    Then $E(\Z) = \{ \O \}$ and
    \begin{equation*}
        E(\Z_p) \neq \{ \O \} \ \Leftrightarrow \ p \neq 3
    \end{equation*}
\end{prop}
\begin{proof}
    First show that there exists some $(x, y) \in \tilde{E}(\F_p)$ with $y \neq 0$ for all primes $p \neq 3$.

@@

    Now consider any prime $p \neq 2, 3$ and $(\tilde{x}, \tilde{y}) \in \tilde{E}(\F_p), x, y \in \Z, \tilde{y} \neq 0$.
    Let $f(t) := t^2 - x^3 - 2x - 2$.
    Then
    \begin{equation*}
        |f(y)|_p \leq p^{-1} \quad \text{and} \quad |f'(y)|_p = |y|_p = 1
    \end{equation*}
    Thus $|f(y)|_p < |f'(y)|_p^2$ and Hensel's Lemma \cite[Thm 2.14]{lecture} yields a root $\gamma \in \Z_p$ with $(x, \gamma) \in E(\Z_p)$.

@@

    For the second part, note that by assumption, there is $(\tilde{x}, \tilde{y}) \in \tilde{E}(\F_p), x, y \in \Z$ with $|y|_p = 1$ and so
    \begin{equation*}
        |y^2 - x^3 - a_2 x^2 - a_4 x - a_6|_p \leq p^{-1} < 1 = 1^2 = |2y|_p
    \end{equation*}
    Hensel's Lemma \cite[Thm 2.14]{lecture} now shows that there is $\gamma \in \Z_p$ such that $\gamma^2 = x^3 + a_2 x^2 + a_4 x + a_6$ and so $(x, \gamma) \in E(\Z_p)$.
    Since $|y|_p = 1$ clearly also $|\gamma|_p = 1$.
    The other direction is obvious and follows directly by taking the reduction modulo $p$.

@@

    So assume now that $p \divides \Delta(E)$, hence $x^3 + f_2 x^2 + f_1 x + f_0$ factors as
    \begin{equation*}
        x^3 + \tilde{f}_2 x^2 + \tilde{f}_1 x + \tilde{f}_0 \equiv (x - \alpha)^2(x - \beta)
    \end{equation*}
    with $\alpha, \beta \in \bar{\F}_p$.
    However, note that $\F_p$ is perfect, so $(x - \alpha)^2(x - \beta)$ cannot be irreducible over $\F_p$, otherwise $\F_p[x] / \langle(x - \alpha)^2(x - \beta)\rangle$ would be a non-separable finite field extension of $\F_p$.
    Thus, either $\alpha \in \F_p$ or $\beta \in \F_p$.
    If $\alpha \in \F_p$, then clearly also $\beta = 2\alpha - \tilde{f}_2 \in \F_p$.
    If $\beta \in \F_p$, then also $(x - \alpha)^2 \in \F_p[x]$ and again by perfectness of $\F_p$, note that $\alpha \in \F_p$.
    So $\alpha, \beta \in \F_p$.

@@

    Interestingly, if we want a curve $E$ such that $\tilde{E}(\F_3) = \{ \O \}$, then we require
    \begin{equation*}
        f(0) = B \equiv 2  \quad \text{and} \quad f(1) = 1 + A + B \equiv 2
    \end{equation*}
    and this has the only solution $A \equiv B \equiv 2$, which indeed works out (this is the curve from Proposition~\ref{prop:1iii}).
\end{example}
This also works for $p = 7$, as the next example shows.
\begin{example}[1(iv) - second example]
    Assume we want to find a curve $E: y^2 = f(x) := x^3 + A x + B$ such that $\tilde{E}(\F_7)$ contains no point $(x, y)$ with $y \neq 0$.
    By considering a scaling isomorphism of the form
    \begin{equation*}
        E \to E', \quad (x, y) \mapsto (u^2 x, u^3 y)
    \end{equation*}
    we can assume wlog that $A \equiv 0, 1, 3 \mod 7$ (as the fourth powers in $\F_7^*$ are $1, 4, 2$).
    Furthermore have
    \begin{equation*}
        f(0) = B \equiv 0, 3, 5, 6, \quad f(1) = 1 + A + B \equiv 0, 3, 5, 6
    \end{equation*}
    since $1, 4, 2$ are the squares in $\F_7^*$.
    So we have the possibilities
    \begin{equation*}
        A \equiv 0, \quad B \equiv 5, 6
    \end{equation*}
    and
    \begin{equation*}
        A \equiv 1, \quad B \equiv 3, 5
    \end{equation*}
    and
    \begin{equation*}
        A \equiv 3, \quad B \equiv 3, 6
    \end{equation*}
    Trying all of them, we find that the only solution such that there is no $(x, y) \in \tilde{E}(\F_7)$ with $y \neq 0$ is $A \equiv 0, B \equiv 6$.
    Indeed, have that
    \begin{equation*}
        \tilde{E}(\F_7) = \{ (1, 0), (2, 0), (4, 0), \O \}
    \end{equation*}
    Furthermore, note that
    \begin{align*}
        2^2 &\equiv 3^3 + 6 = 28 \mod 8 \\
        1^2 &\equiv 1^3 + 6 = 7 \mod 3 \\
        2^2 &\equiv 2^3 + 6 = 14 \mod 5
    \end{align*}
    so Hensel's Lemma \cite[Thm 1.15]{lecture} yields points $(x, y) \in E(\Z_2)$ with $|y|_2 = 1$ and $(x', y') \in E(\Z_3)$ with $|y'|_3 = 1$ and $(x'', y'') \in E(\Z_5)$ with $|y''|_5 = 1$.
\end{example}
\begin{example}[(iv) - third example]
    The cubic polynomial $f = x^3 + x$ has roots $0$ and $1$ modulo 2.
    Thus the Elliptic Curve $E: y^2 = x^3 + x$ clearly has no points $(x, y) \in E(\Z_2)$ with $|y|_2 = 1$, as all points $(x, y) \in \tilde{E}(\F_2)$ satisfy $y = 0$.
    Furthermore, see that
    \begin{equation*}
        \tilde{E}(\F_5) = \{ (0, 0), (2, 0), (3, 0), \O \}
    \end{equation*}
    and
    \begin{align*}
        1^2 &\equiv 2^3 + 2 = 10 \mod 3 \\
        3^2 &\equiv 1^3 + 1 = 2 \mod 7
    \end{align*}
    so Hensel's Lemma yields \cite[Thm 1.15]{lecture} points $(x, y) \in E(\Z_3)$ with $|y|_3 = 1$ and $(x', y') \in E(\Z_7)$ with $|y'|_7 = 1$.
\end{example}
Note also that those three examples show that the bound $p \geq 11$ in Proposition~\ref{prop:points_in_EZp} is sharp.
